### 2018-02-21 Paper-1-Vitis.Rmd
Commit from RStudio

---
title: "Paper 1: Vitis"
output: html_notebook
---

# Intro

The size and shape (physiognomy) of woody, dicotyledonous angiosperm leaves are correlated to climate.  These relationships have been used to develop paleoclimate proxies. Many plant communities, particularly in Northern Hemisphere temperate regions, have an inverse correlation between temperature and the degree of leaf dissection and the presence of teeth, as well as tooth size and abundance. One of the underlying assumptions of leaf physiognomic-paleoclimate proxies is that leaf traits can plastically respond to climate such that that leaves reliably track changes in temperature and precipitation.  Additionally, it is also assumed that leaf traits allometrically scale as a leaf matures through the growing season, such that juvenile and mature leaves have similar leaf traits. However, little work has been done to determine how quickly leaves plastically respond to changes in climate or whether all leaves on a plant respond in the same way.

We used Digital Leaf Physiognomy (DiLP) to measure leaf characters of multiple *Vitis* species from two growing seasons with different temperatures and precipitation. The same vines were sampled each year in order to compare leaf shape on a vine-by-vine and leaf-by-leaf basis.

```{r}
library(tidyverse)
library(tidyr)
library(skimr)
library(readxl)
read_excel("R_DiLP_all_species.xlsx")
excel_sheets("R_DiLP_all_species.xlsx")
read_excel("R_DiLP_Vitis_all_added.xlsx")
excel_sheets("R_DiLP_Vitis_all_added.xlsx")
```

# t-tests
First, we tested whether there were significant differences between the two growing seasons for each of the ratios measured, as well as the total number of teeth. To do this, we ran a series of t-tests.

We first tested each of the variables for *Vitis acerifolia*.

```{r}
ac_all <- read_excel("R_DilP_Vitis_all_added.xlsx", sheet = "acerifolia.all.90")
str(ac_all)
ac_all$year <- as.factor(ac_all$year)
ac_all$tooth.area <- as.numeric(ac_all$tooth.area)
ac_all$tooth.area.perimeter <- as.numeric(ac_all$tooth.area.perimeter)
ac_all$tooth.area.int.perimeter <- as.numeric(ac_all$tooth.area.int.perimeter)
ac_all$average.tooth.area <- as.numeric(ac_all$average.tooth.area)
ac_all$tooth.area.blade.area <- as.numeric(ac_all$tooth.area.blade.area)
skim(ac_all)
t.test(ac_all$total.teeth~ac_all$year, var.equal = TRUE)
t.test(ac_all$feret.diameter.ratio~ac_all$year, var.equal = TRUE)
t.test(ac_all$tooth.area.perimeter~ac_all$year, var.equal = TRUE)
t.test(ac_all$tooth.area.int.perimeter~ac_all$year, var.equal = TRUE)
t.test(ac_all$average.tooth.area~ac_all$year, var.equal = TRUE)
t.test(ac_all$tooth.area.blade.area~ac_all$year, var.equal = TRUE)
t.test(ac_all$teeth.perimeter~ac_all$year, var.equal = TRUE)
t.test(ac_all$teeth.int.perimeter~ac_all$year, var.equal = TRUE)
t.test(ac_all$perimeter.area~ac_all$year, var.equal = TRUE)
t.test(ac_all$perimeter.ratio~ac_all$year, var.equal = TRUE)
t.test(ac_all$compactness~ac_all$year, var.equal = TRUE)
t.test(ac_all$shape.factor~ac_all$year, var.equal = TRUE)
t.test(ac_all$teeth.blade.area~ac_all$year, var.equal = TRUE)
```
Only one variable showed a statistically significant changes between growing seasons: tooth area to blade area (TABA).

We next tested each of the variables for *Vitis aestivalis*.

```{r}
av_all <- read_excel("R_DilP_Vitis_all_added.xlsx", sheet = "aestivalis.all.90")
av_all$year <- as.factor(av_all$year)
av_all$petiole.width.area <- as.numeric(av_all$petiole.width.area)
av_all$petiole.area <- as.numeric(av_all$petiole.area)
av_all$major.feret <- as.numeric(av_all$major.feret)
av_all$tooth.area <- as.numeric(av_all$tooth.area)
av_all$feret.diameter.ratio <- as.numeric(av_all$feret.diameter.ratio)
av_all$tooth.area.perimeter <- as.numeric(av_all$tooth.area.perimeter)
av_all$tooth.area.int.perimeter <- as.numeric(av_all$tooth.area.int.perimeter)
av_all$average.tooth.area <- as.numeric(av_all$average.tooth.area)
av_all$tooth.area.blade.area <- as.numeric(av_all$tooth.area.blade.area)
str(av_all)
skim(av_all)
t.test(av_all$total.teeth~av_all$year, var.equal = TRUE)
t.test(av_all$feret.diameter.ratio~av_all$year, var.equal = TRUE)
t.test(av_all$tooth.area.perimeter~av_all$year, var.equal = TRUE)
t.test(av_all$tooth.area.int.perimeter~av_all$year, var.equal = TRUE)
t.test(av_all$average.tooth.area~av_all$year, var.equal = TRUE)
t.test(av_all$tooth.area.blade.area~av_all$year, var.equal = TRUE)
t.test(av_all$teeth.perimeter~av_all$year, var.equal = TRUE)
t.test(av_all$teeth.int.perimeter~av_all$year, var.equal = TRUE)
t.test(av_all$perimeter.area~av_all$year, var.equal = TRUE)
t.test(av_all$perimeter.ratio~av_all$year, var.equal = TRUE)
t.test(av_all$compactness~av_all$year, var.equal = TRUE)
t.test(av_all$shape.factor~av_all$year, var.equal = TRUE)
t.test(av_all$teeth.blade.area~av_all$year, var.equal = TRUE)
```
Four variables showed statistically significant changes between growing seasons: total teeth (TT), tooth area to blade area (TABA), compactness (CP), shape factor (SF).

We next tested each of the variables for *Vitis amurensis*.

```{r}
am_all <- read_excel("R_DilP_Vitis_all_added.xlsx", sheet = "amurensis.all.90")
am_all$year <- as.factor(am_all$year)
am_all$petiole.width <- as.numeric(am_all$petiole.width)
am_all$petiole.width.area <- as.numeric(am_all$petiole.width.area)
am_all$petiole.area <- as.numeric(am_all$petiole.area)
am_all$blade.area <- as.numeric(am_all$blade.area)
am_all$feret.diameter <- as.numeric(am_all$feret.diameter)
am_all$major.feret <- as.numeric(am_all$major.feret)
am_all$leaf.area <- as.numeric(am_all$leaf.area)
am_all$perimeter <- as.numeric(am_all$perimeter)
am_all$internal.perimeter <- as.numeric(am_all$internal.perimeter)
am_all$feret.diameter.ratio <- as.numeric(am_all$feret.diameter.ratio)
am_all$tooth.area.perimeter <- as.numeric(am_all$tooth.area.perimeter)
am_all$tooth.area.int.perimeter <- as.numeric(am_all$tooth.area.int.perimeter)
am_all$tooth.area.blade.area <- as.numeric(am_all$tooth.area.blade.area)
am_all$teeth.perimeter <- as.numeric(am_all$teeth.perimeter)
am_all$teeth.int.perimeter <- as.numeric(am_all$teeth.int.perimeter)
am_all$perimeter.area <- as.numeric(am_all$perimeter.area)
am_all$perimeter.ratio <- as.numeric(am_all$perimeter.ratio)
am_all$compactness <- as.numeric(am_all$compactness)
am_all$shape.factor <- as.numeric(am_all$shape.factor)
am_all$teeth.blade.area <- as.numeric(am_all$teeth.blade.area)
str(am_all)
skim(am_all)
t.test(am_all$total.teeth~am_all$year, var.equal = TRUE)
t.test(am_all$feret.diameter.ratio~am_all$year, var.equal = TRUE)
t.test(am_all$tooth.area.perimeter~am_all$year, var.equal = TRUE)
t.test(am_all$tooth.area.int.perimeter~am_all$year, var.equal = TRUE)
t.test(am_all$average.tooth.area~am_all$year, var.equal = TRUE)
t.test(am_all$tooth.area.blade.area~am_all$year, var.equal = TRUE)
t.test(am_all$teeth.perimeter~am_all$year, var.equal = TRUE)
t.test(am_all$teeth.int.perimeter~am_all$year, var.equal = TRUE)
t.test(am_all$perimeter.area~am_all$year, var.equal = TRUE)
t.test(am_all$perimeter.ratio~am_all$year, var.equal = TRUE)
t.test(am_all$compactness~am_all$year, var.equal = TRUE)
t.test(am_all$shape.factor~am_all$year, var.equal = TRUE)
t.test(am_all$teeth.blade.area~am_all$year, var.equal = TRUE)
```

Many variables showed statistically significant changes between growing seasons: total teeth (TT), tooth area to perimeter (TAP), tooth area to internal perimeter (TAIP), average tooth area (ATA), *total teeth to perimeter (TP)*, *perimeter to area (PA)*, perimeter to ratio (PR), *compactness (CP)*, and shape factor (SF).

We finally tested each of the variables for *Vitis riparia*.

```{r}
r_all <- read_excel("R_DilP_Vitis_all_added.xlsx", sheet = "riparia.all.90")
r_all$year <- as.factor(r_all$year)
r_all$petiole.width <- as.numeric(r_all$petiole.width)
r_all$petiole.width.area <- as.numeric(r_all$petiole.width.area)
r_all$petiole.area <- as.numeric(r_all$petiole.area)
r_all$internal.perimeter <- as.numeric(r_all$internal.perimeter)
r_all$tooth.area <- as.numeric(r_all$tooth.area)
r_all$tooth.area.perimeter <- as.numeric(r_all$tooth.area.perimeter)
r_all$tooth.area.int.perimeter <- as.numeric(r_all$tooth.area.int.perimeter)
r_all$average.tooth.area <- as.numeric(r_all$average.tooth.area)
r_all$tooth.area.blade.area <- as.numeric(r_all$tooth.area.blade.area)
r_all$teeth.int.perimeter <- as.numeric(r_all$teeth.int.perimeter)
r_all$perimeter.ratio <- as.numeric(r_all$perimeter.ratio)
r_all$teeth.blade.area <- as.numeric(r_all$teeth.blade.area)
str(r_all)
skim(r_all)
t.test(r_all$total.teeth~r_all$year, var.equal = TRUE)
t.test(r_all$feret.diameter.ratio~r_all$year, var.equal = TRUE)
t.test(r_all$tooth.area.perimeter~r_all$year, var.equal = TRUE)
t.test(r_all$tooth.area.int.perimeter~r_all$year, var.equal = TRUE)
t.test(r_all$average.tooth.area~r_all$year, var.equal = TRUE)
t.test(r_all$tooth.area.blade.area~r_all$year, var.equal = TRUE)
t.test(r_all$teeth.perimeter~r_all$year, var.equal = TRUE)
t.test(r_all$teeth.int.perimeter~r_all$year, var.equal = TRUE)
t.test(r_all$perimeter.area~r_all$year, var.equal = TRUE)
t.test(r_all$perimeter.ratio~r_all$year, var.equal = TRUE)
t.test(r_all$compactness~r_all$year, var.equal = TRUE)
t.test(r_all$shape.factor~r_all$year, var.equal = TRUE)
t.test(r_all$teeth.blade.area~r_all$year, var.equal = TRUE)
```

Many variables showed statistically significant changes between growing seasons: tooth area to internal perimeter (TAIP), tooth area to blade area (TABA), total teeth to perimeter (TP), total teeth to internal perimeter (TIP), perimeter ratio (PR), compactness (CP), and shape factor (SF).

# Principal Components Analysis (PCA)

```{r}
library(pcaMethods)
excel_sheets("R_DiLP_Vitis_all_added.xlsx")
```

First, we applied a PCA to *Vitis acerifolia*.
```{r}
ac_ln <- read_excel("R_DiLP_Vitis_all_added.xlsx", sheet = "acerifolia.all.ln")
ac_ln_mat <- data.matrix(ac_ln, rownames.force = NA)
ac_ln_fixed <- ac_ln_mat[,-(1:3)]
ac_ln_fixed[ac_ln_fixed == 0] <- NA
#pca_ac_all <- pca(ac_ln_fixed, method = "nipals", center = TRUE, nPcs = 5)
bpca_ac_ll <- pca(ac_ln_fixed, method = "bpca", nPcs = 5)
summary(bpca_ac_ll)
loadings(bpca_ac_ll)
#summary(pca_ac_all)
#loadings(pca_ac_all)
print(bpca_ac_ll)
#print(pca_ac_all)
#slplot(pca_ac_all)
#loadings_pca_ac <- loadings(pca_ac_all)
#loadings_pca_ac_df <- as.data.frame(loadings_pca_ac)
#str(loadings_pca_ac_df)
#ggplot(loadings_pca_ac_df, aes(x = PC1, y = PC2, label = rownames(loadings_pca_ac_df))) +
  #geom_col(x = PC1, y = PC2) +
  #ggtitle("PCA plot of Vitis acerifolia")
#attach row names
```

We then applied a PCA to *Vitis aestivalis*.

```{r}
av_ln <- read_excel("R_DiLP_Vitis_all_added.xlsx", sheet = "aestivalis.all.ln")
av_ln_mat <- data.matrix(av_ln, rownames.force = NA)
av_ln_fixed <- av_ln_mat[,-(1:3)]
av_ln_fixed[av_ln_fixed == 0] <- NA
#pca_av_all <- pca(av_ln_fixed, method = "nipals", center = TRUE, nPcs = 5)
bpca_av_ll <- pca(av_ln_fixed, method = "bpca", nPcs = 5)
summary(bpca_av_ll)
loadings(bpca_av_ll)
#summary(pca_av_all)
#loadings(pca_av_all)
print(bpca_av_ll)
#print(pca_av_all)
#slplot(pca_av_all)
```

We then applied a PCA to *Vitis amurensis*.

```{r}
am_ln <- read_excel("R_DiLP_Vitis_all_added.xlsx", sheet = "amurensis.all.ln")
am_ln_mat <- data.matrix(am_ln, rownames.force = NA)
am_ln_fixed <- am_ln_mat[,-(1:3)]
am_ln_fixed[am_ln_fixed == 0] <- NA
#pca_am_all <- pca(am_ln_fixed, method = "nipals", center = TRUE, nPcs = 5)
bpca_am_ll <- pca(am_ln_fixed, method = "bpca", nPcs = 5)
summary(bpca_am_ll)
loadings(bpca_am_ll)
#summary(pca_am_all)
#loadings(pca_am_all)
print(bpca_am_ll)
#print(pca_am_all)
#slplot(pca_am_all)
```

Finally, we applied a PCA to *Vitis riparia*.

```{r}
r_ln <- read_excel("R_DiLP_Vitis_all_added.xlsx", sheet = "riparia.all.ln")
r_ln_mat <- data.matrix(r_ln, rownames.force = NA)
r_ln_fixed <- r_ln_mat[,-(1:3)]
r_ln_fixed[r_ln_fixed == 0] <- NA
#pca_r_all <- pca(r_ln_fixed, method = "nipals", center = TRUE, nPcs = 5)
bpca_r_ll <- pca(r_ln_fixed, method = "bpca", nPcs = 5)
summary(bpca_r_ll)
loadings(bpca_r_ll)
#summary(pca_r_all)
#loadings(pca_r_all)
print(bpca_r_ll)
#print(pca_r_all)
#slplot(pca_r_all)
```

# Generalized Linear Modeling

First, we tested *Vitis acerifolia*.

We ran a GLM for leaf number and total teeth.

```{r}
skim(ac_all)
ac_tt_glm <- glm(leaf~total.teeth, data = ac_all, family = gaussian())
summary(ac_tt_glm)
plot(ac_tt_glm)
```

We ran a GLM for leaf number and feret diameter ratio.

```{r}
ac_fdr_glm <- glm(leaf~feret.diameter.ratio, data = ac_all, family = gaussian())
summary(ac_fdr_glm)
plot(ac_fdr_glm)
```

We ran a GLM for leaf number and tooth area:perimeter.

```{r}
ac_tap_glm <- glm(leaf~tooth.area.perimeter, data = ac_all, family = gaussian())
summary(ac_tap_glm)
plot(ac_tap_glm)
```

We ran a GLM for leaf number and tooth area:internal perimeter.

```{r}
ac_taip_glm <- glm(leaf~tooth.area.int.perimeter, data = ac_all, family = gaussian())
summary(ac_taip_glm)
plot(ac_taip_glm)
```

We ran a GLM for leaf number and average tooth area.

```{r}
ac_ata_glm <- glm(leaf~average.tooth.area, data = ac_all, family = gaussian())
summary(ac_ata_glm)
plot(ac_ata_glm)
```

We ran a GLM for leaf number and tooth area:blade area..

```{r}
ac_taba_glm <- glm(leaf~tooth.area.blade.area, data = ac_all, family = gaussian())
summary(ac_taba_glm)
plot(ac_taba_glm)
```

We ran a GLM for leaf number and teeth:perimeter.

```{r}
ac_tp_glm <- glm(leaf~teeth.perimeter, data = ac_all, family = gaussian())
summary(ac_tp_glm)
plot(ac_tp_glm)
```

We ran a GLM for leaf number and teeth:internal perimeter.

```{r}
ac_tip_glm <- glm(leaf~teeth.int.perimeter, data = ac_all, family = gaussian())
summary(ac_tip_glm)
plot(ac_tip_glm)
```

We ran a GLM for leaf number and perimeter:area.

```{r}
ac_pa_glm <- glm(leaf~perimeter.area, data = ac_all, family = gaussian())
summary(ac_pa_glm)
plot(ac_pa_glm)
```

We ran a GLM for leaf number and perimeter ratio.

```{r}
ac_pr_glm <- glm(leaf~perimeter.ratio, data = ac_all, family = gaussian())
summary(ac_pr_glm)
plot(ac_pr_glm)
```

We ran a GLM for leaf number and compactness.

```{r}
ac_cp_glm <- glm(leaf~compactness, data = ac_all, family = gaussian())
summary(ac_cp_glm)
plot(ac_cp_glm)
```

We ran a GLM for leaf number and shape factor.

```{r}
ac_sf_glm <- glm(leaf~shape.factor, data = ac_all, family = gaussian())
summary(ac_sf_glm)
plot(ac_sf_glm)
```

We ran a GLM for leaf number and teeth:blade area.

```{r}
ac_tba_glm <- glm(leaf~teeth.blade.area, data = ac_all, family = gaussian())
summary(ac_tba_glm)
plot(ac_tba_glm)
```

We ran a GLM for leaf number and multiple variables. 
```{r}
ac_clim_glm <- glm(leaf~feret.diameter.ratio + tooth.area.blade.area + teeth.blade.area + teeth.int.perimeter, data = ac_all, family = gaussian())
summary(ac_clim_glm)
plot(ac_clim_glm)
```

We ran a GLM for leaf number and all variables. 
```{r}
ac_all_glm <- glm(leaf~total.teeth + feret.diameter.ratio + tooth.area.perimeter + tooth.area.int.perimeter + average.tooth.area + tooth.area.blade.area + teeth.perimeter + teeth.int.perimeter + perimeter.area + perimeter.ratio + compactness + shape.factor + teeth.blade.area, data = ac_all, family = gaussian())
summary(ac_all_glm)
plot(ac_all_glm)
```

We tested *Vitis aestivalis*.

We ran a GLM for leaf number and total teeth.

```{r}
skim(av_all)
av_tt_glm <- glm(leaf~total.teeth, data = av_all, family = gaussian())
summary(av_tt_glm)
plot(av_tt_glm)
```

We ran a GLM for leaf number and feret diameter ratio.

```{r}
av_fdr_glm <- glm(leaf~feret.diameter.ratio, data = av_all, family = gaussian())
summary(av_fdr_glm)
plot(av_fdr_glm)
```

We ran a GLM for leaf number and tooth area:perimeter.

```{r}
av_tap_glm <- glm(leaf~tooth.area.perimeter, data = av_all, family = gaussian())
summary(av_tap_glm)
plot(av_tap_glm)
```

We ran a GLM for leaf number and tooth area:internal perimeter.

```{r}
av_taip_glm <- glm(leaf~tooth.area.int.perimeter, data = av_all, family = gaussian())
summary(av_taip_glm)
plot(av_taip_glm)
```

We ran a GLM for leaf number and average tooth area.

```{r}
av_ata_glm <- glm(leaf~average.tooth.area, data = av_all, family = gaussian())
summary(av_ata_glm)
plot(av_ata_glm)
```

We ran a GLM for leaf number and tooth area:blade area..

```{r}
av_taba_glm <- glm(leaf~tooth.area.blade.area, data = av_all, family = gaussian())
summary(av_taba_glm)
plot(av_taba_glm)
```

We ran a GLM for leaf number and teeth:perimeter.

```{r}
av_tp_glm <- glm(leaf~teeth.perimeter, data = av_all, family = gaussian())
summary(av_tp_glm)
plot(av_tp_glm)
```

We ran a GLM for leaf number and teeth:internal perimeter.

```{r}
av_tip_glm <- glm(leaf~teeth.int.perimeter, data = av_all, family = gaussian())
summary(av_tip_glm)
plot(av_tip_glm)
```

We ran a GLM for leaf number and perimeter:area.

```{r}
av_pa_glm <- glm(leaf~perimeter.area, data = av_all, family = gaussian())
summary(av_pa_glm)
plot(av_pa_glm)
```

We ran a GLM for leaf number and perimeter ratio.

```{r}
av_pr_glm <- glm(leaf~perimeter.ratio, data = av_all, family = gaussian())
summary(av_pr_glm)
plot(av_pr_glm)
```

We ran a GLM for leaf number and compactness.

```{r}
av_cp_glm <- glm(leaf~compactness, data = av_all, family = gaussian())
summary(av_cp_glm)
plot(av_cp_glm)
```

We ran a GLM for leaf number and shape factor.

```{r}
av_sf_glm <- glm(leaf~shape.factor, data = av_all, family = gaussian())
summary(av_sf_glm)
plot(av_sf_glm)
```

We ran a GLM for leaf number and teeth:blade area.

```{r}
av_tba_glm <- glm(leaf~teeth.blade.area, data = av_all, family = gaussian())
summary(av_tba_glm)
plot(av_tba_glm)
```

We ran a GLM for leaf number and multiple variables. 
```{r}
av_clim_glm <- glm(leaf~feret.diameter.ratio + tooth.area.blade.area + teeth.blade.area + teeth.int.perimeter, data = av_all, family = gaussian())
summary(av_clim_glm)
plot(av_clim_glm)
```

We ran a GLM for leaf number and all variables. 
```{r}
av_all_glm <- glm(leaf~total.teeth + feret.diameter.ratio + tooth.area.perimeter + tooth.area.int.perimeter + average.tooth.area + tooth.area.blade.area + teeth.perimeter + teeth.int.perimeter + perimeter.area + perimeter.ratio + compactness + shape.factor + teeth.blade.area, data = av_all, family = gaussian())
summary(av_all_glm)
plot(av_all_glm)
```

we tested *Vitis amurensis*.

We ran a GLM for leaf number and total teeth.

```{r}
skim(am_all)
am_tt_glm <- glm(leaf~total.teeth, data = am_all, family = gaussian())
summary(am_tt_glm)
plot(am_tt_glm)
```

We ran a GLM for leaf number and feret diameter ratio.

```{r}
am_fdr_glm <- glm(leaf~feret.diameter.ratio, data = am_all, family = gaussian())
summary(am_fdr_glm)
plot(am_fdr_glm)
```

We ran a GLM for leaf number and tooth area:perimeter.

```{r}
am_tap_glm <- glm(leaf~tooth.area.perimeter, data = am_all, family = gaussian())
summary(am_tap_glm)
plot(am_tap_glm)
```

We ran a GLM for leaf number and tooth area:internal perimeter.

```{r}
am_taip_glm <- glm(leaf~tooth.area.int.perimeter, data = am_all, family = gaussian())
summary(am_taip_glm)
plot(am_taip_glm)
```

We ran a GLM for leaf number and average tooth area.

```{r}
am_ata_glm <- glm(leaf~average.tooth.area, data = am_all, family = gaussian())
summary(am_ata_glm)
plot(am_ata_glm)
```

We ran a GLM for leaf number and tooth area:blade area..

```{r}
am_taba_glm <- glm(leaf~tooth.area.blade.area, data = am_all, family = gaussian())
summary(am_taba_glm)
plot(am_taba_glm)
```

We ran a GLM for leaf number and teeth:perimeter.

```{r}
am_tp_glm <- glm(leaf~teeth.perimeter, data = am_all, family = gaussian())
summary(am_tp_glm)
plot(am_tp_glm)
```

We ran a GLM for leaf number and teeth:internal perimeter.

```{r}
am_tip_glm <- glm(leaf~teeth.int.perimeter, data = am_all, family = gaussian())
summary(am_tip_glm)
plot(am_tip_glm)
```

We ran a GLM for leaf number and perimeter:area.

```{r}
am_pa_glm <- glm(leaf~perimeter.area, data = am_all, family = gaussian())
summary(am_pa_glm)
plot(am_pa_glm)
```

We ran a GLM for leaf number and perimeter ratio.

```{r}
am_pr_glm <- glm(leaf~perimeter.ratio, data = am_all, family = gaussian())
summary(am_pr_glm)
plot(am_pr_glm)
```

We ran a GLM for leaf number and compactness.

```{r}
am_cp_glm <- glm(leaf~compactness, data = am_all, family = gaussian())
summary(am_cp_glm)
plot(am_cp_glm)
```

We ran a GLM for leaf number and shape factor.

```{r}
am_sf_glm <- glm(leaf~shape.factor, data = am_all, family = gaussian())
summary(am_sf_glm)
plot(am_sf_glm)
```

We ran a GLM for leaf number and teeth:blade area.

```{r}
am_tba_glm <- glm(leaf~teeth.blade.area, data = am_all, family = gaussian())
summary(am_tba_glm)
plot(am_tba_glm)
```

We ran a GLM for leaf number and multiple variables. 
```{r}
am_clim_glm <- glm(leaf~feret.diameter.ratio + tooth.area.blade.area + teeth.blade.area + teeth.int.perimeter, data = am_all, family = gaussian())
summary(am_clim_glm)
plot(am_clim_glm)
```

We ran a GLM for leaf number and all variables. 
```{r}
am_all_glm <- glm(leaf~total.teeth + feret.diameter.ratio + tooth.area.perimeter + tooth.area.int.perimeter + average.tooth.area + tooth.area.blade.area + teeth.perimeter + teeth.int.perimeter + perimeter.area + perimeter.ratio + compactness + shape.factor + teeth.blade.area, data = am_all, family = gaussian())
summary(am_all_glm)
plot(am_all_glm)
```

We tested *Vitis riparia*.

We ran a GLM for leaf number and total teeth.

```{r}
skim(r_all)
r_tt_glm <- glm(leaf~total.teeth, data = r_all, family = gaussian())
summary(r_tt_glm)
plot(r_tt_glm)
```

We ran a GLM for leaf number and feret diameter ratio.

```{r}
r_fdr_glm <- glm(leaf~feret.diameter.ratio, data = r_all, family = gaussian())
summary(r_fdr_glm)
plot(r_fdr_glm)
```

We ran a GLM for leaf number and tooth area:perimeter.

```{r}
r_tap_glm <- glm(leaf~tooth.area.perimeter, data = r_all, family = gaussian())
summary(r_tap_glm)
plot(r_tap_glm)
```

We ran a GLM for leaf number and tooth area:internal perimeter.

```{r}
r_taip_glm <- glm(leaf~tooth.area.int.perimeter, data = r_all, family = gaussian())
summary(r_taip_glm)
plot(r_taip_glm)
```

We ran a GLM for leaf number and average tooth area.

```{r}
r_ata_glm <- glm(leaf~average.tooth.area, data = r_all, family = gaussian())
summary(r_ata_glm)
plot(r_ata_glm)
```

We ran a GLM for leaf number and tooth area:blade area.

```{r}
r_taba_glm <- glm(leaf~tooth.area.blade.area, data = r_all, family = gaussian())
summary(r_taba_glm)
plot(r_taba_glm)
```

We ran a GLM for leaf number and teeth:perimeter.

```{r}
r_tp_glm <- glm(leaf~teeth.perimeter, data = r_all, family = gaussian())
summary(r_tp_glm)
plot(r_tp_glm)
```

We ran a GLM for leaf number and teeth:internal perimeter.

```{r}
r_tip_glm <- glm(leaf~teeth.int.perimeter, data = r_all, family = gaussian())
summary(r_tip_glm)
plot(r_tip_glm)
```

We ran a GLM for leaf number and perimeter:area.

```{r}
r_pa_glm <- glm(leaf~perimeter.area, data = r_all, family = gaussian())
summary(r_pa_glm)
plot(r_pa_glm)
```

We ran a GLM for leaf number and perimeter ratio.

```{r}
r_pr_glm <- glm(leaf~perimeter.ratio, data = r_all, family = gaussian())
summary(r_pr_glm)
plot(r_pr_glm)
```

We ran a GLM for leaf number and compactness.

```{r}
r_cp_glm <- glm(leaf~compactness, data = r_all, family = gaussian())
summary(r_cp_glm)
plot(r_cp_glm)
```

We ran a GLM for leaf number and shape factor.

```{r}
r_sf_glm <- glm(leaf~shape.factor, data = r_all, family = gaussian())
summary(r_sf_glm)
plot(r_sf_glm)
```

We ran a GLM for leaf number and teeth:blade area.

```{r}
r_tba_glm <- glm(leaf~teeth.blade.area, data = r_all, family = gaussian())
summary(r_tba_glm)
plot(r_tba_glm)
```

We ran a GLM for leaf number and multiple variables. 
```{r}
r_clim_glm <- glm(leaf~feret.diameter.ratio + tooth.area.blade.area + teeth.blade.area + teeth.int.perimeter, data = r_all, family = gaussian())
summary(r_clim_glm)
plot(r_clim_glm)
```

We ran a GLM for leaf number and all variables. 
```{r}
r_all_glm <- glm(leaf~total.teeth + feret.diameter.ratio + tooth.area.perimeter + tooth.area.int.perimeter + average.tooth.area + tooth.area.blade.area + teeth.perimeter + teeth.int.perimeter + perimeter.area + perimeter.ratio + compactness + shape.factor + teeth.blade.area, data = r_all, family = gaussian())
summary(r_all_glm)
plot(r_all_glm)
beep(sound = 3, expr = NULL)
```