### 2018-10-29 V_amurensis_LM.Rmd
V. amurensis prep

---
title: "Vitis amurensis LM"
output: html_notebook
---

# Intro

The size and shape (physiognomy) of woody, dicotyledonous angiosperm leaves are correlated to climate.  These relationships have been used to develop paleoclimate proxies. Many plant communities, particularly in Northern Hemisphere temperate regions, have an inverse correlation between temperature and the degree of leaf dissection and the presence of teeth, as well as tooth size and abundance. One of the underlying assumptions of leaf physiognomic-paleoclimate proxies is that leaf traits can plastically respond to climate such that that leaves reliably track changes in temperature and precipitation.  Additionally, it is also assumed that leaf traits allometrically scale as a leaf matures through the growing season, such that juvenile and mature leaves have similar leaf traits. However, little work has been done to determine how quickly leaves plastically respond to changes in climate or whether all leaves on a plant respond in the same way.

We used Digital Leaf Physiognomy (DiLP) to measure leaf characters of multiple *Vitis* species from two growing seasons with different temperatures and precipitation. The same vines were sampled each year in order to compare leaf shape on a vine-by-vine and leaf-by-leaf basis.

```{r setup}
library(readxl)
library(dplyr)
library(skimr)
library(ggplot2)
library(gganimate)
library(segmented)
Vitis_90_ratios <- read_excel("R_DiLP_all_species.xlsx", 
     sheet = "all.90", col_types = c("text", 
         "numeric", "numeric", "numeric", 
         "skip", "skip", "skip", "skip", 
         "skip", "skip", "skip", "skip", 
         "skip", "skip", "skip", "numeric", 
         "skip", "numeric", "numeric", "numeric", 
         "numeric", "numeric", "numeric", 
         "numeric", "numeric", "numeric", 
         "numeric", "numeric", "numeric"))
amurensis <-filter(Vitis_90_ratios, species == "amurensis")
```

# t-tests
First, we tested whether there were significant differences between the two growing seasons for each of the ratios measured, as well as the total number of teeth. To do this, we ran a series of t-tests.

We next tested each of the variables for *Vitis amurensis*.

```{r}
t.test(total.teeth ~ year, data= amurensis, var.equal= TRUE)
t.test(feret.diameter.ratio ~ year, data= amurensis, var.equal= TRUE)
t.test(tooth.area.perimeter ~ year, data= amurensis, var.equal= TRUE)
t.test(tooth.area.int.perimeter ~ year, data= amurensis, var.equal= TRUE)
t.test(average.tooth.area ~ year, data= amurensis, var.equal= TRUE)
t.test(tooth.area.blade.area ~ year, data= amurensis, var.equal= TRUE)
t.test(teeth.perimeter ~ year, data= amurensis, var.equal= TRUE)
t.test(teeth.int.perimeter ~ year, data= amurensis, var.equal= TRUE)
t.test(perimeter.area ~ year, data= amurensis, var.equal= TRUE)
t.test(perimeter.ratio ~ year, data= amurensis, var.equal= TRUE)
t.test(compactness ~ year, data= amurensis, var.equal= TRUE)
t.test(shape.factor ~ year, data= amurensis, var.equal= TRUE)
t.test(teeth.blade.area ~ year, data= amurensis, var.equal= TRUE)
```
Many variables showed statistically significant changes between growing seasons: total teeth (TT), tooth area to perimeter (TAP), tooth area to internal perimeter (TAIP), average tooth area (ATA), perimeter to ratio (PR), and shape factor (SF).

# Linear Modeling

Next we tested *Vitis amurensis*.

We need the means for each leaf character for both years combined.
```{r}
am_l_means <-  amurensis %>% 
  group_by(leaf) %>% 
  summarize(avg.tt = mean(total.teeth, na.rm = TRUE),
            avg.fdr = mean(feret.diameter.ratio, na.rm = TRUE),
            avg.tap = mean(tooth.area.perimeter, na.rm = TRUE),
            avg.taip = mean(tooth.area.int.perimeter, na.rm = TRUE),
            avg.ta = mean(average.tooth.area, na.rm = TRUE),
            avg.taba = mean(tooth.area.blade.area, na.rm = TRUE),
            avg.tp = mean(teeth.perimeter, na.rm = TRUE),
            avg.tip = mean(teeth.int.perimeter, na.rm = TRUE),
            avg.pa = mean(perimeter.area, na.rm = TRUE),
            avg.pr = mean(perimeter.ratio, na.rm = TRUE),
            avg.cp = mean(compactness, na.rm = TRUE),
            avg.sf = mean(shape.factor, na.rm = TRUE),
            avg.tba = mean(teeth.blade.area, na.rm = TRUE))
```

Many variables showed statistically significant changes between growing seasons: total teeth (TT), tooth area to perimeter (TAP), tooth area to internal perimeter (TAIP), average tooth area (ATA), perimeter to ratio (PR), and shape factor (SF), so we need to split these variables for each year.
```{r}
am_13 <- filter(amurensis, year == 2013)
am_15 <- filter(amurensis, year == 2015)
```

Total teeth
```{r}
am_tt13_mean <-  amurensis %>% 
  filter(year == 2013) %>% 
  group_by(leaf) %>% 
  summarize(avg.tt = mean(total.teeth, na.rm = TRUE))

am_tt15_mean <- amurensis %>% 
  filter(year == 2015) %>% 
  group_by(leaf) %>% 
  summarize(avg.tt = mean(total.teeth, na.rm = TRUE))
```

Tooth area:perimeter
```{r}
am_tap13_mean <-  amurensis %>% 
  filter(year == 2013) %>% 
  group_by(leaf) %>% 
  summarize(avg.tap = mean(tooth.area.perimeter, na.rm = TRUE))

am_tap15_mean <- amurensis %>% 
  filter(year == 2015) %>% 
  group_by(leaf) %>% 
  summarize(avg.tap = mean(tooth.area.perimeter, na.rm = TRUE))
```

Tooth area:internal perimeter
```{r}
am_taip13_mean <-  amurensis %>% 
  filter(year == 2013) %>% 
  group_by(leaf) %>% 
  summarize(avg.taip = mean(tooth.area.int.perimeter, na.rm = TRUE))

am_taip15_mean <- amurensis %>% 
  filter(year == 2015) %>% 
  group_by(leaf) %>% 
  summarize(avg.taip = mean(tooth.area.int.perimeter, na.rm = TRUE))
```

Average tooth area
```{r}
am_ata13_mean <-  amurensis %>% 
  filter(year == 2013) %>% 
  group_by(leaf) %>% 
  summarize(avg.ata = mean(average.tooth.area, na.rm = TRUE))

am_ata15_mean <- amurensis %>% 
  filter(year == 2015) %>% 
  group_by(leaf) %>% 
  summarize(avg.ata = mean(average.tooth.area, na.rm = TRUE))
```

Perimeter ratio
```{r}
am_pr13_mean <- amurensis %>% 
  filter(year == 2013) %>% 
  group_by(leaf) %>% 
  summarize(avg.pr = mean(perimeter.ratio, na.rm = TRUE))

am_pr15_mean <- amurensis %>% 
  filter(year == 2015) %>% 
  group_by(leaf) %>% 
  summarize(avg.pr = mean(perimeter.ratio, na.rm = TRUE))
```

Shape factor
```{r}
am_sf13_mean <- amurensis %>% 
  filter(year == 2013) %>% 
  group_by(leaf) %>% 
  summarize(avg.sf = mean(shape.factor, na.rm = TRUE))

am_sf15_mean <- amurensis %>% 
  filter(year == 2015) %>% 
  group_by(leaf) %>% 
  summarize(avg.sf = mean(shape.factor, na.rm = TRUE))
```

We tested leaf number and feret diameter ratio.
```{r}
am_fdr_plot <- ggplot(amurensis, aes(x = leaf, y = feret.diameter.ratio)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 13, 1))
am_fdr_plot <- am_fdr_plot +labs(x = "Leaf number",
                               y = "FDR")
am_fdr_plot

```

```{r}
am_fdr_mean_plot <- ggplot(am_l_means, aes(x = leaf, y = avg.fdr)) + geom_line() + theme_classic()
am_fdr_mean_plot <- am_fdr_mean_plot +labs(x = "Leaf number",
                               y = "FDR")
am_fdr_mean_plot
```

We ran a LM for leaf number and feret diameter ratio.
```{r}
am_fdr_lm <- lm(avg.fdr ~ leaf, data = am_l_means)
summary(am_fdr_lm)
```

```{r}
am_fdr_seg <- segmented(am_fdr_lm,
                       seg.Z = ~ leaf,
                       psi = list(leaf = c(4, 7, 11)))
summary(am_fdr_seg)
```

```{r}
am_fdr_seg$psi
slope(am_fdr_seg)
am_fdr_fitted <- fitted(am_fdr_seg)
am_fdr_model <- data.frame(Leaf = am_l_means$leaf, FDR = am_fdr_fitted)
ggplot(am_fdr_model, aes(x = Leaf, y = FDR)) + geom_line()
am_fdr_lines <- am_fdr_seg$psi[, 2]
am_fdr_mean_break <- am_fdr_plot + geom_vline(xintercept = am_fdr_lines, linetype = "dashed")
am_fdr_mean_break
ggsave("am_fdr_break.png")
```