### 2018-11-02 V_amurensis_LM.Rmd
V. amurensis complete

---
title: "Vitis amurensis LM"
output: html_notebook
---

# Intro

The size and shape (physiognomy) of woody, dicotyledonous angiosperm leaves are correlated to climate.  These relationships have been used to develop paleoclimate proxies. Many plant communities, particularly in Northern Hemisphere temperate regions, have an inverse correlation between temperature and the degree of leaf dissection and the presence of teeth, as well as tooth size and abundance. One of the underlying assumptions of leaf physiognomic-paleoclimate proxies is that leaf traits can plastically respond to climate such that that leaves reliably track changes in temperature and precipitation.  Additionally, it is also assumed that leaf traits allometrically scale as a leaf matures through the growing season, such that juvenile and mature leaves have similar leaf traits. However, little work has been done to determine how quickly leaves plastically respond to changes in climate or whether all leaves on a plant respond in the same way.

We used Digital Leaf Physiognomy (DiLP) to measure leaf characters of multiple *Vitis* species from two growing seasons with different temperatures and precipitation. The same vines were sampled each year in order to compare leaf shape on a vine-by-vine and leaf-by-leaf basis.

```{r setup}
library(readxl)
library(dplyr)
library(skimr)
library(ggplot2)
library(gganimate)
library(segmented)
Vitis_90_ratios <- read_excel("R_DiLP_all_species.xlsx", 
     sheet = "all.90", col_types = c("text", 
         "numeric", "numeric", "numeric", 
         "skip", "skip", "skip", "skip", 
         "skip", "skip", "skip", "skip", 
         "skip", "skip", "skip", "numeric", 
         "skip", "numeric", "numeric", "numeric", 
         "numeric", "numeric", "numeric", 
         "numeric", "numeric", "numeric", 
         "numeric", "numeric", "numeric"))
amurensis <-filter(Vitis_90_ratios, species == "amurensis")
```

# t-tests
First, we tested whether there were significant differences between the two growing seasons for each of the ratios measured, as well as the total number of teeth. To do this, we ran a series of t-tests.

We next tested each of the variables for *Vitis amurensis*.

```{r}
t.test(total.teeth ~ year, data= amurensis, var.equal= TRUE)
t.test(feret.diameter.ratio ~ year, data= amurensis, var.equal= TRUE)
t.test(tooth.area.perimeter ~ year, data= amurensis, var.equal= TRUE)
t.test(tooth.area.int.perimeter ~ year, data= amurensis, var.equal= TRUE)
t.test(average.tooth.area ~ year, data= amurensis, var.equal= TRUE)
t.test(tooth.area.blade.area ~ year, data= amurensis, var.equal= TRUE)
t.test(teeth.perimeter ~ year, data= amurensis, var.equal= TRUE)
t.test(teeth.int.perimeter ~ year, data= amurensis, var.equal= TRUE)
t.test(perimeter.area ~ year, data= amurensis, var.equal= TRUE)
t.test(perimeter.ratio ~ year, data= amurensis, var.equal= TRUE)
t.test(compactness ~ year, data= amurensis, var.equal= TRUE)
t.test(shape.factor ~ year, data= amurensis, var.equal= TRUE)
t.test(teeth.blade.area ~ year, data= amurensis, var.equal= TRUE)
```
Many variables showed statistically significant changes between growing seasons: total teeth (TT), tooth area to perimeter (TAP), tooth area to internal perimeter (TAIP), average tooth area (ATA), perimeter to ratio (PR), and shape factor (SF).

# Linear Modeling

Next we tested *Vitis amurensis*.

We need the means for each leaf character for both years combined.
```{r}
am_l_means <-  amurensis %>% 
  group_by(leaf) %>% 
  summarize(avg.tt = mean(total.teeth, na.rm = TRUE),
            avg.fdr = mean(feret.diameter.ratio, na.rm = TRUE),
            avg.tap = mean(tooth.area.perimeter, na.rm = TRUE),
            avg.taip = mean(tooth.area.int.perimeter, na.rm = TRUE),
            avg.ta = mean(average.tooth.area, na.rm = TRUE),
            avg.taba = mean(tooth.area.blade.area, na.rm = TRUE),
            avg.tp = mean(teeth.perimeter, na.rm = TRUE),
            avg.tip = mean(teeth.int.perimeter, na.rm = TRUE),
            avg.pa = mean(perimeter.area, na.rm = TRUE),
            avg.pr = mean(perimeter.ratio, na.rm = TRUE),
            avg.cp = mean(compactness, na.rm = TRUE),
            avg.sf = mean(shape.factor, na.rm = TRUE),
            avg.tba = mean(teeth.blade.area, na.rm = TRUE))
```

Many variables showed statistically significant changes between growing seasons: total teeth (TT), tooth area to perimeter (TAP), tooth area to internal perimeter (TAIP), average tooth area (ATA), perimeter to ratio (PR), and shape factor (SF), so we need to split these variables for each year.
```{r}
am_13 <- filter(amurensis, year == 2013)
am_15 <- filter(amurensis, year == 2015)
```

Total teeth
```{r}
am_tt13_mean <-  amurensis %>% 
  filter(year == 2013) %>% 
  group_by(leaf) %>% 
  summarize(avg.tt = mean(total.teeth, na.rm = TRUE))

am_tt15_mean <- amurensis %>% 
  filter(year == 2015) %>% 
  group_by(leaf) %>% 
  summarize(avg.tt = mean(total.teeth, na.rm = TRUE))
```

Tooth area:perimeter
```{r}
am_tap13_mean <-  amurensis %>% 
  filter(year == 2013) %>% 
  group_by(leaf) %>% 
  summarize(avg.tap = mean(tooth.area.perimeter, na.rm = TRUE))

am_tap15_mean <- amurensis %>% 
  filter(year == 2015) %>% 
  group_by(leaf) %>% 
  summarize(avg.tap = mean(tooth.area.perimeter, na.rm = TRUE))
```

Tooth area:internal perimeter
```{r}
am_taip13_mean <-  amurensis %>% 
  filter(year == 2013) %>% 
  group_by(leaf) %>% 
  summarize(avg.taip = mean(tooth.area.int.perimeter, na.rm = TRUE))

am_taip15_mean <- amurensis %>% 
  filter(year == 2015) %>% 
  group_by(leaf) %>% 
  summarize(avg.taip = mean(tooth.area.int.perimeter, na.rm = TRUE))
```

Average tooth area
```{r}
am_ata13_mean <-  amurensis %>% 
  filter(year == 2013) %>% 
  group_by(leaf) %>% 
  summarize(avg.ata = mean(average.tooth.area, na.rm = TRUE))

am_ata15_mean <- amurensis %>% 
  filter(year == 2015) %>% 
  group_by(leaf) %>% 
  summarize(avg.ata = mean(average.tooth.area, na.rm = TRUE))
```

Perimeter ratio
```{r}
am_pr13_mean <- amurensis %>% 
  filter(year == 2013) %>% 
  group_by(leaf) %>% 
  summarize(avg.pr = mean(perimeter.ratio, na.rm = TRUE))

am_pr15_mean <- amurensis %>% 
  filter(year == 2015) %>% 
  group_by(leaf) %>% 
  summarize(avg.pr = mean(perimeter.ratio, na.rm = TRUE))
```

Shape factor
```{r}
am_sf13_mean <- amurensis %>% 
  filter(year == 2013) %>% 
  group_by(leaf) %>% 
  summarize(avg.sf = mean(shape.factor, na.rm = TRUE))

am_sf15_mean <- amurensis %>% 
  filter(year == 2015) %>% 
  group_by(leaf) %>% 
  summarize(avg.sf = mean(shape.factor, na.rm = TRUE))
```

We tested leaf number and feret diameter ratio.
```{r}
am_fdr_plot <- ggplot(amurensis, aes(x = leaf, y = feret.diameter.ratio)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 13, 1))
am_fdr_plot <- am_fdr_plot +labs(x = "Leaf number",
                               y = "FDR")
am_fdr_plot

```

```{r}
am_fdr_mean_plot <- ggplot(am_l_means, aes(x = leaf, y = avg.fdr)) + geom_line() + theme_classic()
am_fdr_mean_plot <- am_fdr_mean_plot +labs(x = "Leaf number",
                               y = "FDR")
am_fdr_mean_plot
```

We ran a LM for leaf number and feret diameter ratio.
```{r}
am_fdr_lm <- lm(avg.fdr ~ leaf, data = am_l_means)
summary(am_fdr_lm)
```

```{r}
am_fdr_seg <- segmented(am_fdr_lm,
                       seg.Z = ~ leaf,
                       psi = list(leaf = c(3, 7, 10)))
summary(am_fdr_seg)
```

```{r}
am_fdr_seg$psi
slope(am_fdr_seg)
am_fdr_fitted <- fitted(am_fdr_seg)
am_fdr_model <- data.frame(Leaf = am_l_means$leaf, FDR = am_fdr_fitted)
ggplot(am_fdr_model, aes(x = Leaf, y = FDR)) + geom_line()
am_fdr_lines <- am_fdr_seg$psi[, 2]
am_fdr_mean_break <- am_fdr_plot + geom_vline(xintercept = am_fdr_lines, linetype = "dashed")
am_fdr_mean_break
ggsave("am_fdr_break.png")
```

We tested leaf number and tooth area:blade area.
```{r}
am_taba_plot <- ggplot(amurensis, aes(x = leaf, y = tooth.area.blade.area)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 13, 1))
am_taba_plot <- am_taba_plot +labs(x = "Leaf number",
                               y = "Tooth area:blade area")
am_taba_plot

```

```{r}
am_taba_mean_plot <- ggplot(am_l_means, aes(x = leaf, y = avg.taba)) + geom_line() + theme_classic()
am_taba_mean_plot <- am_taba_mean_plot +labs(x = "Leaf number",
                               y = "Tooth area:blade area")
am_taba_mean_plot
```

We ran a LM for leaf number and tooth area:blade area.
```{r}
am_taba_lm <- lm(avg.taba ~ leaf, data = am_l_means)
summary(am_taba_lm)
```

```{r}
am_taba_seg <- segmented(am_taba_lm,
                       seg.Z = ~ leaf,
                       psi = list(leaf = c(3, 8, 11)))
summary(am_taba_seg)
```

```{r}
am_taba_seg$psi
slope(am_taba_seg)
am_taba_fitted <- fitted(am_taba_seg)
am_taba_model <- data.frame(Leaf = am_l_means$leaf, TABA = am_taba_fitted)
ggplot(am_taba_model, aes(x = Leaf, y = TABA)) + geom_line()
am_taba_lines <- am_taba_seg$psi[, 2]
am_taba_mean_break <- am_taba_plot + geom_vline(xintercept = am_taba_lines, linetype = "dashed")
am_taba_mean_break
ggsave("am_taba_break.png")
```

We tested leaf number and total teeth:perimeter.
```{r}
am_tp_plot <- ggplot(amurensis, aes(x = leaf, y = teeth.perimeter)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 13, 1))
am_tp_plot <- am_tp_plot +labs(x = "Leaf number",
                               y = "Teeth:perimeter")
am_tp_plot

```

```{r}
am_tp_mean_plot <- ggplot(am_l_means, aes(x = leaf, y = avg.tp)) + geom_line() + theme_classic()
am_tp_mean_plot <- am_tp_mean_plot +labs(x = "Leaf number",
                               y = "Teeth.perimeter")
am_tp_mean_plot
```

We ran a LM for leaf number and teeth:perimeter.
```{r}
am_tp_lm <- lm(avg.tp ~ leaf, data = am_l_means)
summary(am_tp_lm)
```

```{r}
am_tp_seg <- segmented(am_tp_lm,
                       seg.Z = ~ leaf,
                       psi = list(leaf = c(3, 8, 11)))
summary(am_tp_seg)
```

```{r}
am_tp_seg$psi
slope(am_tp_seg)
am_tp_fitted <- fitted(am_tp_seg)
am_tp_model <- data.frame(Leaf = am_l_means$leaf, TP = am_tp_fitted)
ggplot(am_tp_model, aes(x = Leaf, y = TP)) + geom_line()
am_tp_lines <- am_tp_seg$psi[, 2]
am_tp_mean_break <- am_tp_plot + geom_vline(xintercept = am_tp_lines, linetype = "dashed")
am_tp_mean_break
ggsave("am_tp_break.png")
```

We tested leaf number and total teeth:internal perimeter.
```{r}
am_tip_plot <- ggplot(amurensis, aes(x = leaf, y = teeth.int.perimeter)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 13, 1))
am_tip_plot <- am_tip_plot +labs(x = "Leaf number",
                               y = "Teeth:internal perimeter")
am_tip_plot
```

```{r}
am_tip_mean_plot <- ggplot(am_l_means, aes(x = leaf, y = avg.tip)) + geom_line() + theme_classic()
am_tip_mean_plot <- am_tip_mean_plot +labs(x = "Leaf number",
                               y = "Teeth.internal perimeter")
am_tip_mean_plot
```

We ran a LM for leaf number and teeth:internal perimeter.
```{r}
am_tip_lm <- lm(avg.tip ~ leaf, data = am_l_means)
summary(am_tip_lm)
```

```{r}
am_tip_seg <- segmented(am_tip_lm,
                       seg.Z = ~ leaf,
                       psi = list(leaf = c(3, 8, 11)))
summary(am_tip_seg)
```

```{r}
am_tip_seg$psi
slope(am_tip_seg)
am_tip_fitted <- fitted(am_tip_seg)
am_tip_model <- data.frame(Leaf = am_l_means$leaf, TIP = am_tip_fitted)
ggplot(am_tip_model, aes(x = Leaf, y = TIP)) + geom_line()
am_tip_lines <- am_tip_seg$psi[, 2]
am_tip_mean_break <- am_tip_plot + geom_vline(xintercept = am_tip_lines, linetype = "dashed")
am_tip_mean_break
ggsave("am_tip_break.png")
```

We tested leaf number and perimeter:area.
```{r}
am_pa_plot <- ggplot(amurensis, aes(x = leaf, y = perimeter.area)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 13, 1))
am_pa_plot <- am_pa_plot +labs(x = "Leaf number",
                               y = "Perimeter:area")
am_pa_plot
```

```{r}
am_pa_mean_plot <- ggplot(am_l_means, aes(x = leaf, y = avg.pa)) + geom_line() + theme_classic()
am_pa_mean_plot <- am_pa_mean_plot +labs(x = "Leaf number",
                               y = "Perimeter:area")
am_pa_mean_plot
```

We ran a LM for leaf number and perimeter:area.
```{r}
am_pa_lm <- lm(avg.pa ~ leaf, data = am_l_means)
summary(am_pa_lm)
```

```{r}
am_pa_seg <- segmented(am_pa_lm,
                       seg.Z = ~ leaf,
                       psi = list(leaf = c(3, 8, 11)))
summary(am_pa_seg)
```

```{r}
am_pa_seg$psi
slope(am_pa_seg)
am_pa_fitted <- fitted(am_pa_seg)
am_pa_model <- data.frame(Leaf = am_l_means$leaf, PA = am_pa_fitted)
ggplot(am_pa_model, aes(x = Leaf, y = PA)) + geom_line()
am_pa_lines <- am_pa_seg$psi[, 2]
am_pa_mean_break <- am_pa_plot + geom_vline(xintercept = am_pa_lines, linetype = "dashed")
am_pa_mean_break
ggsave("am_pa_break.png")
```

We tested leaf number and compactness.
```{r}
am_cp_plot <- ggplot(amurensis, aes(x = leaf, y = compactness)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 13, 1))
am_cp_plot <- am_cp_plot +labs(x = "Leaf number",
                               y = "Compactness")
am_cp_plot
```

```{r}
am_cp_mean_plot <- ggplot(am_l_means, aes(x = leaf, y = avg.cp)) + geom_line() + theme_classic()
am_cp_mean_plot <- am_cp_mean_plot +labs(x = "Leaf number",
                               y = "Compactness")
am_cp_mean_plot
```

We ran a LM for leaf number and compactness.
```{r}
am_cp_lm <- lm(avg.cp ~ leaf, data = am_l_means)
summary(am_cp_lm)
```

```{r}
am_cp_seg <- segmented(am_cp_lm,
                       seg.Z = ~ leaf,
                       psi = list(leaf = c(3, 8, 11)))
summary(am_cp_seg)
```

```{r}
am_cp_seg$psi
slope(am_cp_seg)
am_cp_fitted <- fitted(am_cp_seg)
am_cp_model <- data.frame(Leaf = am_l_means$leaf, CP = am_cp_fitted)
ggplot(am_cp_model, aes(x = Leaf, y = CP)) + geom_line()
am_cp_lines <- am_cp_seg$psi[, 2]
am_cp_mean_break <- am_cp_plot + geom_vline(xintercept = am_cp_lines, linetype = "dashed")
am_cp_mean_break
ggsave("am_cp_break.png")
```

We tested leaf number and total teeth:blade are.
```{r}
am_tba_plot <- ggplot(amurensis, aes(x = leaf, y = teeth.blade.area)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 13, 1))
am_tba_plot <- am_tba_plot +labs(x = "Leaf number",
                               y = "Total Teeth:Blade Area")
am_tba_plot
```

```{r}
am_tba_mean_plot <- ggplot(am_l_means, aes(x = leaf, y = avg.tba)) + geom_line() + theme_classic()
am_tba_mean_plot <- am_tba_mean_plot +labs(x = "Leaf number",
                               y = "Total Teeth:Blade Area")
am_tba_mean_plot
```

We ran a LM for leaf number and total teeth:blade area.
```{r}
am_tba_lm <- lm(avg.tba ~ leaf, data = am_l_means)
summary(am_tba_lm)
```

```{r}
am_tba_seg <- segmented(am_tba_lm,
                       seg.Z = ~ leaf,
                       psi = list(leaf = c(3, 8, 11)))
summary(am_tba_seg)
```

```{r}
am_tba_seg$psi
slope(am_tba_seg)
am_tba_fitted <- fitted(am_tba_seg)
am_tba_model <- data.frame(Leaf = am_l_means$leaf, TBA = am_tba_fitted)
ggplot(am_tba_model, aes(x = Leaf, y = TBA)) + geom_line()
am_tba_lines <- am_tba_seg$psi[, 2]
am_tba_mean_break <- am_tba_plot + geom_vline(xintercept = am_tba_lines, linetype = "dashed")
am_tba_mean_break
ggsave("am_tba_break.png")
```

We tested leaf number and total teeth for 2013.
```{r}
am_tt13_plot <- ggplot(am_13, aes(x = leaf, y = total.teeth)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 12, 1))
am_tt13_plot <- am_tt13_plot +labs(x = "Leaf number",
                               y = "Total Teeth")
am_tt13_plot
```

```{r}
am_tt13_mean_plot <- ggplot(am_tt13_mean, aes(x = leaf, y = avg.tt)) + geom_line() + theme_classic()
am_tt13_mean_plot <- am_tt13_mean_plot +labs(x = "Leaf number",
                               y = "Total Teeth")
am_tt13_mean_plot
```

We ran a LM for leaf number and total teeth in 2013.
```{r}
am_tt13_lm <- lm(avg.tt ~ leaf, data = am_tt13_mean)
summary(am_tt13_lm)
```

```{r}
am_tt13_seg <- segmented(am_tt13_lm,
                       seg.Z = ~ leaf,
                       psi = list(leaf = c(3, 7, 10)))
summary(am_tt13_seg)
```

```{r}
am_tt13_seg$psi
slope(am_tt13_seg)
am_tt13_fitted <- fitted(am_tt13_seg)
am_tt13_model <- data.frame(Leaf = am_tt13_mean$leaf, TT = am_tt13_fitted)
ggplot(am_tt13_model, aes(x = Leaf, y = TT)) + geom_line()
am_tt13_lines <- am_tt13_seg$psi[, 2]
am_tt13_mean_break <- am_tt13_plot + geom_vline(xintercept = am_tt13_lines, linetype = "dashed")
am_tt13_mean_break
ggsave("am_tt13_break.png")
```

We tested leaf number and total teeth for 2015.
```{r}
am_tt15_plot <- ggplot(am_15, aes(x = leaf, y = total.teeth)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 12, 1))
am_tt15_plot <- am_tt15_plot +labs(x = "Leaf number",
                               y = "Total Teeth")
am_tt15_plot
```

```{r}
am_tt15_mean_plot <- ggplot(am_tt15_mean, aes(x = leaf, y = avg.tt)) + geom_line() + theme_classic()
am_tt15_mean_plot <- am_tt15_mean_plot +labs(x = "Leaf number",
                               y = "Total Teeth")
am_tt15_mean_plot
```

We ran a LM for leaf number and total teeth in 2015.
```{r}
am_tt15_lm <- lm(avg.tt ~ leaf, data = am_tt15_mean)
summary(am_tt15_lm)
```

```{r}
am_tt15_seg <- segmented(am_tt15_lm,
                       seg.Z = ~ leaf,
                       psi = list(leaf = c(3, 7, 10)))
summary(am_tt15_seg)
```

```{r}
am_tt15_seg$psi
slope(am_tt15_seg)
am_tt15_fitted <- fitted(am_tt15_seg)
am_tt15_model <- data.frame(Leaf = am_tt15_mean$leaf, TT = am_tt15_fitted)
ggplot(am_tt15_model, aes(x = Leaf, y = TT)) + geom_line()
am_tt15_lines <- am_tt15_seg$psi[, 2]
am_tt15_mean_break <- am_tt15_plot + geom_vline(xintercept = am_tt15_lines, linetype = "dashed")
am_tt15_mean_break
ggsave("am_tt15_break.png")
```

We tested leaf number and tooth area:perimeter for 2013.
```{r}
am_tap13_plot <- ggplot(am_13, aes(x = leaf, y = tooth.area.perimeter)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 12, 1))
am_tap13_plot <- am_tap13_plot +labs(x = "Leaf number",
                               y = "Tooth area: perimeter")
am_tap13_plot
```

```{r}
am_tap13_mean_plot <- ggplot(am_tap13_mean, aes(x = leaf, y = avg.tap)) + geom_line() + theme_classic()
am_tap13_mean_plot <- am_tap13_mean_plot +labs(x = "Leaf number",
                               y = "Tooth area: perimeter")
am_tap13_mean_plot
```

We ran a LM for leaf number and toth area perimeter in 2013.
```{r}
am_tap13_lm <- lm(avg.tap ~ leaf, data = am_tap13_mean)
summary(am_tap13_lm)
```

```{r}
am_tap13_seg <- segmented(am_tap13_lm,
                       seg.Z = ~ leaf,
                       psi = list(leaf = c(3, 7, 10)))
summary(am_tap13_seg)
```

```{r}
am_tap13_seg$psi
slope(am_tap13_seg)
am_tap13_fitted <- fitted(am_tap13_seg)
am_tap13_model <- data.frame(Leaf = am_tap13_mean$leaf, TAP = am_tap13_fitted)
ggplot(am_tap13_model, aes(x = Leaf, y = TAP)) + geom_line()
am_tap13_lines <- am_tap13_seg$psi[, 2]
am_tap13_mean_break <- am_tap13_plot + geom_vline(xintercept = am_tap13_lines, linetype = "dashed")
am_tap13_mean_break
ggsave("am_tap13_break.png")
```

We tested leaf number and tooth area:perimeter for 2015.
```{r}
am_tap15_plot <- ggplot(am_15, aes(x = leaf, y = tooth.area.perimeter)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 12, 1))
am_tap15_plot <- am_tap15_plot +labs(x = "Leaf number",
                               y = "Tooth area:perimeter")
am_tap15_plot
```

```{r}
am_tap15_mean_plot <- ggplot(am_tap15_mean, aes(x = leaf, y = avg.tap)) + geom_line() + theme_classic()
am_tap15_mean_plot <- am_tap15_mean_plot +labs(x = "Leaf number",
                               y = "Tooth area:perimeter")
am_tap15_mean_plot
```

We ran a LM for leaf number and tooth area:perimeter in 2015.
```{r}
am_tap15_lm <- lm(avg.tap ~ leaf, data = am_tap15_mean)
summary(am_tap15_lm)
```

```{r}
am_tap15_seg <- segmented(am_tap15_lm,
                       seg.Z = ~ leaf,
                       psi = list(leaf = c(3, 7, 10)))
summary(am_tap15_seg)
```

```{r}
am_tap15_seg$psi
slope(am_tap15_seg)
am_tap15_fitted <- fitted(am_tap15_seg)
am_tap15_model <- data.frame(Leaf = am_tap15_mean$leaf, TAP = am_tap15_fitted)
ggplot(am_tap15_model, aes(x = Leaf, y = TAP)) + geom_line()
am_tap15_lines <- am_tap15_seg$psi[, 2]
am_tap15_mean_break <- am_tap15_plot + geom_vline(xintercept = am_tap15_lines, linetype = "dashed")
am_tap15_mean_break
ggsave("am_tap15_break.png")
```

We tested leaf number and tooth area:internal perimeter for 2013.
```{r}
am_taip13_plot <- ggplot(am_13, aes(x = leaf, y = tooth.area.int.perimeter)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 12, 1))
am_taip13_plot <- am_taip13_plot +labs(x = "Leaf number",
                               y = "Tooth area: internal perimeter")
am_taip13_plot
```

```{r}
am_taip13_mean_plot <- ggplot(am_taip13_mean, aes(x = leaf, y = avg.taip)) + geom_line() + theme_classic()
am_taip13_mean_plot <- am_taip13_mean_plot +labs(x = "Leaf number",
                               y = "Tooth area:internal perimeter")
am_taip13_mean_plot
```

We ran a LM for leaf number and toth area: internal perimeter in 2013.
```{r}
am_taip13_lm <- lm(avg.taip ~ leaf, data = am_taip13_mean)
summary(am_taip13_lm)
```

```{r}
am_taip13_seg <- segmented(am_taip13_lm,
                       seg.Z = ~ leaf,
                       psi = list(leaf = c(3, 7, 10)))
summary(am_taip13_seg)
```

```{r}
am_taip13_seg$psi
slope(am_taip13_seg)
am_taip13_fitted <- fitted(am_taip13_seg)
am_taip13_model <- data.frame(Leaf = am_taip13_mean$leaf, TAIP = am_taip13_fitted)
ggplot(am_taip13_model, aes(x = Leaf, y = TAIP)) + geom_line()
am_taip13_lines <- am_taip13_seg$psi[, 2]
am_taip13_mean_break <- am_taip13_plot + geom_vline(xintercept = am_taip13_lines, linetype = "dashed")
am_taip13_mean_break
ggsave("am_taip13_break.png")
```

We tested leaf number and tooth area:internal perimeter for 2015.
```{r}
am_taip15_plot <- ggplot(am_15, aes(x = leaf, y = tooth.area.int.perimeter)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 12, 1))
am_taip15_plot <- am_taip15_plot +labs(x = "Leaf number",
                               y = "Tooth area:internal perimeter")
am_taip15_plot
```

```{r}
am_taip15_mean_plot <- ggplot(am_taip15_mean, aes(x = leaf, y = avg.taip)) + geom_line() + theme_classic()
am_taip15_mean_plot <- am_taip15_mean_plot +labs(x = "Leaf number",
                               y = "Tooth area:internal perimeter")
am_taip15_mean_plot
```

We ran a LM for leaf number and tooth area:internal perimeter in 2015.
```{r}
am_taip15_lm <- lm(avg.taip ~ leaf, data = am_taip15_mean)
summary(am_taip15_lm)
```

```{r}
am_taip15_seg <- segmented(am_taip15_lm,
                       seg.Z = ~ leaf,
                       psi = list(leaf = c(3, 7, 10)))
summary(am_taip15_seg)
```

```{r}
am_taip15_seg$psi
slope(am_taip15_seg)
am_taip15_fitted <- fitted(am_taip15_seg)
am_taip15_model <- data.frame(Leaf = am_taip15_mean$leaf, TAIP = am_taip15_fitted)
ggplot(am_taip15_model, aes(x = Leaf, y = TAIP)) + geom_line()
am_taip15_lines <- am_taip15_seg$psi[, 2]
am_taip15_mean_break <- am_taip15_plot + geom_vline(xintercept = am_taip15_lines, linetype = "dashed")
am_taip15_mean_break
ggsave("am_taip15_break.png")
```

We tested leaf number and average tooth area for 2013.
```{r}
am_ata13_plot <- ggplot(am_13, aes(x = leaf, y = average.tooth.area)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 12, 1))
am_ata13_plot <- am_ata13_plot +labs(x = "Leaf number",
                               y = "Average tooth area")
am_ata13_plot
```

```{r}
am_ata13_mean_plot <- ggplot(am_ata13_mean, aes(x = leaf, y = avg.ata)) + geom_line() + theme_classic()
am_ata13_mean_plot <- am_ata13_mean_plot +labs(x = "Leaf number",
                               y = "Average tooth area")
am_ata13_mean_plot
```

We ran a LM for leaf number and average tooth area in 2013.
```{r}
am_ata13_lm <- lm(avg.ata ~ leaf, data = am_ata13_mean)
summary(am_ata13_lm)
```

```{r}
am_ata13_seg <- segmented(am_ata13_lm,
                       seg.Z = ~ leaf,
                       psi = list(leaf = c(3, 7, 10)))
summary(am_ata13_seg)
```

```{r}
am_ata13_seg$psi
slope(am_ata13_seg)
am_ata13_fitted <- fitted(am_ata13_seg)
am_ata13_model <- data.frame(Leaf = am_ata13_mean$leaf, ATA = am_ata13_fitted)
ggplot(am_ata13_model, aes(x = Leaf, y = ATA)) + geom_line()
am_ata13_lines <- am_ata13_seg$psi[, 2]
am_ata13_mean_break <- am_ata13_plot + geom_vline(xintercept = am_ata13_lines, linetype = "dashed")
am_ata13_mean_break
ggsave("am_ata13_break.png")
```

We tested leaf number and average tooth area for 2015.
```{r}
am_ata15_plot <- ggplot(am_15, aes(x = leaf, y = average.tooth.area)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 12, 1))
am_ata15_plot <- am_ata15_plot +labs(x = "Leaf number",
                               y = "Average tooth area")
am_ata15_plot
```

```{r}
am_ata15_mean_plot <- ggplot(am_ata15_mean, aes(x = leaf, y = avg.ata)) + geom_line() + theme_classic()
am_ata15_mean_plot <- am_ata15_mean_plot +labs(x = "Leaf number",
                               y = "Average tooth area")
am_ata15_mean_plot
```

We ran a LM for leaf number and average tooth area in 2015.
```{r}
am_ata15_lm <- lm(avg.ata ~ leaf, data = am_ata15_mean)
summary(am_ata15_lm)
```

```{r}
am_ata15_seg <- segmented(am_ata15_lm,
                       seg.Z = ~ leaf,
                       psi = list(leaf = c(3, 7, 10)))
summary(am_ata15_seg)
```

```{r}
am_ata15_seg$psi
slope(am_ata15_seg)
am_ata15_fitted <- fitted(am_ata15_seg)
am_ata15_model <- data.frame(Leaf = am_ata15_mean$leaf, ATA = am_ata15_fitted)
ggplot(am_ata15_model, aes(x = Leaf, y = ATA)) + geom_line()
am_ata15_lines <- am_ata15_seg$psi[, 2]
am_ata15_mean_break <- am_ata15_plot + geom_vline(xintercept = am_ata15_lines, linetype = "dashed")
am_ata15_mean_break
ggsave("am_ata15_break.png")
```

We tested leaf number and perimeter ratio for 2013.
```{r}
am_pr13_plot <- ggplot(am_13, aes(x = leaf, y = perimeter.ratio)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 12, 1))
am_pr13_plot <- am_pr13_plot +labs(x = "Leaf number",
                               y = "Perimeter ratio")
am_pr13_plot
```

```{r}
am_pr13_mean_plot <- ggplot(am_pr13_mean, aes(x = leaf, y = avg.pr)) + geom_line() + theme_classic()
am_pr13_mean_plot <- am_pr13_mean_plot +labs(x = "Leaf number",
                               y = "Perimeter ratio")
am_pr13_mean_plot
```

We ran a LM for leaf number and perimeter ratio in 2013.
```{r}
am_pr13_lm <- lm(avg.pr ~ leaf, data = am_pr13_mean)
summary(am_pr13_lm)
```

```{r}
am_pr13_seg <- segmented(am_pr13_lm,
                       seg.Z = ~ leaf,
                       psi = list(leaf = c(3, 7, 10)))
summary(am_pr13_seg)
```

```{r}
am_pr13_seg$psi
slope(am_pr13_seg)
am_pr13_fitted <- fitted(am_pr13_seg)
am_pr13_model <- data.frame(Leaf = am_pr13_mean$leaf, PR = am_pr13_fitted)
ggplot(am_pr13_model, aes(x = Leaf, y = PR)) + geom_line()
am_pr13_lines <- am_pr13_seg$psi[, 2]
am_pr13_mean_break <- am_pr13_plot + geom_vline(xintercept = am_pr13_lines, linetype = "dashed")
am_pr13_mean_break
ggsave("am_pr13_break.png")
```

We tested leaf number and perimeter ratio for 2015.
```{r}
am_pr15_plot <- ggplot(am_15, aes(x = leaf, y = perimeter.ratio)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 12, 1))
am_pr15_plot <- am_pr15_plot +labs(x = "Leaf number",
                               y = "Perimeter ratio")
am_pr15_plot
```

```{r}
am_pr15_mean_plot <- ggplot(am_pr15_mean, aes(x = leaf, y = avg.pr)) + geom_line() + theme_classic()
am_pr15_mean_plot <- am_pr15_mean_plot +labs(x = "Leaf number",
                               y = "Perimeter ratio")
am_pr15_mean_plot
```

We ran a LM for leaf number and average tooth area in 2015.
```{r}
am_pr15_lm <- lm(avg.pr ~ leaf, data = am_pr15_mean)
summary(am_pr15_lm)
```

```{r}
am_pr15_seg <- segmented(am_pr15_lm,
                       seg.Z = ~ leaf,
                       psi = list(leaf = c(3, 7, 10)))
summary(am_pr15_seg)
```

```{r}
am_pr15_seg$psi
slope(am_pr15_seg)
am_pr15_fitted <- fitted(am_pr15_seg)
am_pr15_model <- data.frame(Leaf = am_pr15_mean$leaf, PR = am_pr15_fitted)
ggplot(am_pr15_model, aes(x = Leaf, y = PR)) + geom_line()
am_pr15_lines <- am_pr15_seg$psi[, 2]
am_pr15_mean_break <- am_pr15_plot + geom_vline(xintercept = am_pr15_lines, linetype = "dashed")
am_pr15_mean_break
ggsave("am_pr15_break.png")
```

We tested leaf number and shape factor for 2013.
```{r}
am_sf13_plot <- ggplot(am_13, aes(x = leaf, y = shape.factor)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 12, 1))
am_sf13_plot <- am_sf13_plot +labs(x = "Leaf number",
                               y = "Shape factor")
am_sf13_plot
```

```{r}
am_sf13_mean_plot <- ggplot(am_sf13_mean, aes(x = leaf, y = avg.sf)) + geom_line() + theme_classic()
am_sf13_mean_plot <- am_sf13_mean_plot +labs(x = "Leaf number",
                               y = "Shape factor")
am_sf13_mean_plot
```

We ran a LM for leaf number and shape factor in 2013.
```{r}
am_sf13_lm <- lm(avg.sf ~ leaf, data = am_sf13_mean)
summary(am_sf13_lm)
```

```{r}
am_sf13_seg <- segmented(am_sf13_lm,
                       seg.Z = ~ leaf,
                       psi = list(leaf = c(3, 7, 10)))
summary(am_sf13_seg)
```

```{r}
am_sf13_seg$psi
slope(am_sf13_seg)
am_sf13_fitted <- fitted(am_sf13_seg)
am_sf13_model <- data.frame(Leaf = am_sf13_mean$leaf, SF = am_sf13_fitted)
ggplot(am_sf13_model, aes(x = Leaf, y = SF)) + geom_line()
am_sf13_lines <- am_sf13_seg$psi[, 2]
am_sf13_mean_break <- am_sf13_plot + geom_vline(xintercept = am_sf13_lines, linetype = "dashed")
am_sf13_mean_break
ggsave("am_sf13_break.png")
```

We tested leaf number and shape factor for 2015.
```{r}
am_sf15_plot <- ggplot(am_15, aes(x = leaf, y = shape.factor)) + geom_smooth() + theme_classic() + scale_x_continuous(breaks = seq(0, 12, 1))
am_sf15_plot <- am_sf15_plot +labs(x = "Leaf number",
                               y = "Shape factor")
am_sf15_plot
```

```{r}
am_sf15_mean_plot <- ggplot(am_sf15_mean, aes(x = leaf, y = avg.sf)) + geom_line() + theme_classic()
am_sf15_mean_plot <- am_sf15_mean_plot +labs(x = "Leaf number",
                               y = "Shape factor")
am_sf15_mean_plot
```

We ran a LM for leaf number and shape factor in 2015.
```{r}
am_sf15_lm <- lm(avg.sf ~ leaf, data = am_sf15_mean)
summary(am_sf15_lm)
```

```{r}
am_sf15_seg <- segmented(am_sf15_lm,
                       seg.Z = ~ leaf,
                       psi = list(leaf = c(3, 7, 10)))
summary(am_sf15_seg)
```

```{r}
am_sf15_seg$psi
slope(am_sf15_seg)
am_sf15_fitted <- fitted(am_sf15_seg)
am_sf15_model <- data.frame(Leaf = am_sf15_mean$leaf, SF = am_ata15_fitted)
ggplot(am_sf15_model, aes(x = Leaf, y = SF)) + geom_line()
am_sf15_lines <- am_sf15_seg$psi[, 2]
am_sf15_mean_break <- am_sf15_plot + geom_vline(xintercept = am_sf15_lines, linetype = "dashed")
am_sf15_mean_break
ggsave("am_sf15_break.png")
```